Stage-Splitting a Modal Language

### 2014-08-29 paper/techreport.tex
Add introduction to mini-report.

\begin{abstract}
\input{abstract}
\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

@@

\input{intro}

\input{semantics}

\vfill\eject\pagebreak
\input{splitting}
\input{algoexamples}
\input{relatedwork}

@@

\bibliography{paper}
\bibliographystyle{abbrvnat}

>>>>

%\begin{abstract}
%\input{abstract}
%\end{abstract}

%\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

%\keywords
%keyword1, keyword2

@@

\section{Introduction}

This document is a tech report on ongoing research into a program transformation
technique which we call {\em splitting}.  The splitting operation statically transforms a program written in a modal language into multiple unstaged (monostage) terms.

The idea of splitting has previously been explored as an optimization technique that improves performance in situations where multi-variate functions are invoked many times, with some inputs being invariant across invocations and others being dynamic (varying across invocations). Prior examples of splitting simple languages include the ``data specialization" work of Knoblock and Ruff\,\cite{knoblock96}, as well as the compilation of high-level shading languages for computer graphics such as RTSL\,\cite{Proudfoot:2001} and Spark\,\cite{Foley:2011}.  In this document we define the splitting transformation for the two-stage language \lang, a typed lambda calculus featuring products, sums, and isorecursive types.

\begin{itemize}
\item \Cref{sec:semantics} describes \lang which unambiguously specifies the inputs to 
our splitting transformation.  The type system of \lang\ is drawn from prior work, but its semantics are uniquely defined in anticipation of splitting.

\item \Cref{sec:splitting} describes the splitting transformation on \lang.  As stated above, splitting accepts terms in \lang\ as input, and converts them to terms in a standard unstaged language.  This section also formally defines the correctness for the splitting transformation.

\item \Cref{sec:examples} gives several examples of how splitting terms in \lang\ can be used
to derive data structures that are communicated between the two unstaged terms, and discusses the effectiveness and limitations of this technique.  Unlike prior work, where the data communicated between the resultant unstaged terms was a basic tuple of values, splitting \lang\ terms can generate non-trivial intermediate structures, such as trees.
\end{itemize}

\input{semantics}
\input{splitting}
\input{algoexamples}

@@

\bibliography{techreport}
\bibliographystyle{abbrvnat}
